\section*{B.Sc. I Year Second Semester}

(First Week)
Groups, Subgroups and their examples, the group $\Z_n$ of integers under addition modulo ${n}$, The group $U(n)$ of units under multiplication modulo $n$, \\
(Second Week)Cyclic groups, complex roots of unity, circle group, the
general linear group $GL_n (n,R)$, Dihedral group. The commutator subgroup, Examples of subgroups
including the center of a group, Cosets, Index of subgroup, Lagrange Theorem, order of an element.
Normal subgroups: their definition, examples, and characterizations, Quotient groups, Class equations.


Books Recommended

1. John B. Fraleigh, A First Course in Abstract Algebra, 7th Ed., Pearson, 2002.

2. M. Artin, Abstract Algebra, 2nd Ed., Pearson, 2011.

3. Joseph A Gallian, Contemporary Abstract Algebra, 4th Ed., Narosa, 1999.

4. I. Herrnstein, Topics in Algebra, 2nd ed., John Wiley, 1999

\section*{M.Sc. First Year Second Semester}

Review: Group, Subgroups, Normal subgroups, Lagrange Theorem, Isomorphism Theorems, Cayley’s Theorem, Class Equation, Sylow's Theorems. (15 Lectures)

\noindent Ring Theory: Rings, Examples (including formal power series
ring and matrix ring), Sub-ring, Ideals, Prime and Maximal Ideals,
Rings of Fractions, Integral domains, Homomorphism of Rings
and its basic theorems. Polynomial Rings, Gaussian Rings, Euclidean Domains, Principal Ideal Domains and Unique Factorizations, Gauss’ Lemma, Irreducibility Criteria. (25 Lectures)

\noindent Field Theory: Field, Field extension, Algebraic extension, Finite
fields. (15 Lectures)

Books Recommended

[1] I. Herrnstein, Topics in Algebra, 2nd ed., John Wiley, 1999

[2] D. S. Dummit, R. M. Foote, Abstract Algebra, 3rd edition,
Wiley India. 2014.

[3] J. B. Fralieigh, A first course in abstract algebra, 5th ed.,
Addison-Wesley, 1994.

[4] Michael Artin, Algebra, Prentice Hall of India (1991).

[5] Serge Lang, Algebra, revised 3
rd edition, Springer (2004).

[6] J. A. Gallian, Contemporary Abstract Algebra, Narosa.

